Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-10 intro.tex
putting it in Phys Rev E Format
more PRE formatting
more pre formatting
update hopcroft citation

\section{Introduction {\color{blue} EDITABLE}}\label{sec:intro}

%****NOTES START HERE****

@@

As a result computer performance traces are very
complicated and range across the entire spectrum of complexity, making this an ideal test bed.  Even a simple ``microkernel," like a three-line loop that
repeatedly initializes the upper triangle of a matrix in column-major order, can produce everything from periodic to
chaotic performance traces \cite{mytkowicz09}, depending on the architecture running the software. Such a performance trace is shown in
Figure~\ref{fig:ipc}.%, and chaos places fundamental limits on predictability.
%
 \begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{figs/colshortts}
    % where an .eps filename suffix will be assumed under latex,
    % and a .pdf suffix will be assumed for pdflatex
    \caption{A small snippet of the instructions per cycle(ipc) of \col, a three-line C program that repeatedly initializes
      a matrix in column-major order, running on an Intel i7\textsuperscript{\textregistered}-based machine.}
   \label{fig:ipc}
  \end{figure}
%
More interesting programs, e.g., compilers and numerical solvers, produce time series which exist throughout the spectrum of complexity: from completely predictable to completely unstructured,i.e., low to high complexity[[Just trying to reinforce the wording in the readers mind]]. In addition, some performance traces even exhibit interesting regime changes as a program switches between varying subroutines. With a single system (computer performance) capable of producing complicated time series that vary so greatly in complexity (periodic to noise), it would be invaluable to known a priori if a trace contained enough predictive structure to build a  forecast model based on structure before time and effort was spent searching for an ideal model.

@@

% paragraph to appease the theoretician in me
It is worth taking a moment to consider the theoretical possibility of
this task. We are not attempting to predict the state of the CPU at an
arbitrary point in the future --- this, at least with perfect
accuracy, would be tantamount to solving the halting problem. What we
are attempting is to predict aspects or functions of the running of
the CPU, e.g., processor efficiency and similar statistics. Prediction of these quantities
at some finite time in the future, even with perfect accuracy, does
not violate the Rice-Shapiro theorem~\cite{hopcroft} which essentially states that any non-trivial property of a program is uncomputable.

>>>>

\section{Introduction  }\label{sec:intro}
{\color{blue}[[EDITABLE]]}
%****NOTES START HERE****

@@

As a result computer performance traces are very
complicated and range across the entire spectrum of complexity, making this an ideal test bed.  Even a simple ``microkernel," like a three-line loop that
repeatedly initializes the upper triangle of a matrix in column-major order, can produce everything from periodic to
chaotic performance traces \cite{mytkowicz09}, depending on the architecture running the software. Such a performance trace is shown in
Figure~\ref{fig:ipc}.%, and chaos places fundamental limits on predictability.
%
 \begin{figure}[htbp]
    \centering
    \includegraphics[width=\columnwidth]{figs/colshortts}
    % where an .eps filename suffix will be assumed under latex,
    % and a .pdf suffix will be assumed for pdflatex
    \caption{A small snippet of the instructions per cycle(ipc) of \col, a three-line C program that repeatedly initializes
      a matrix in column-major order, running on an Intel i7\textsuperscript{\textregistered}-based machine.}
   \label{fig:ipc}
  \end{figure}
%
More interesting programs, e.g., compilers and numerical solvers, produce time series which exist throughout the spectrum of complexity: from completely predictable to completely unstructured,i.e., low to high complexity[[Just trying to reinforce the wording in the readers mind]]. In addition, some performance traces even exhibit interesting regime changes as a program switches between varying subroutines. With a single system (computer performance) capable of producing complicated time series that vary so greatly in complexity (periodic to noise), it would be invaluable to known a priori if a trace contained enough predictive structure to build a  forecast model based on structure before time and effort was spent searching for an ideal model.

@@

% paragraph to appease the theoretician in me
It is worth taking a moment to consider the theoretical possibility of
this task. We are not attempting to predict the state of the CPU at an
arbitrary point in the future --- this, at least with perfect
accuracy, would be tantamount to solving the halting problem. What we
are attempting is to predict aspects or functions of the running of
the CPU, e.g., processor efficiency and similar statistics. Prediction of these quantities
at some finite time in the future, even with perfect accuracy, does
not violate the Rice-Shapiro theorem~\cite{hopcroft2007} which essentially states that any non-trivial property of a program is uncomputable.
